---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
weather_1 <- read_csv("data/delhi_weather_one.csv")
weather_2 <- read_csv("data/delhi_weather_two.csv")
```

###Create a function to store calculations
```{r}
#Find average temp, humidity and windspeed ----

weather_report <- function(weather){
  # Our average temperature, humidity and wind speed 
  weather_means <- 
  weather %>% 
  summarise(
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_windspeed = mean(wind_speed)
  )

# Show hot and humid days
hot_and_humid <- 
weather %>% 
  filter(temp > 20) %>% 
  filter(humidity > 55)

return(list(weather_means, hot_and_humid))
}

weather_report(weather_1)
```

###Use the above function and apply it to second dataset
This runs with no errors
_but_
it does return some odd values in 2nd table
```{r}
weather_2 %>% 
  weather_report()
```

###Using assertr

this can help throw _errors_ in your code
used when you want to use a range of _expected_ outputs 

```{r}
library(assertr)
```

###We can give a range of inputs to our weather data

temp: -35 - 55
humidity: 0 - 100
wind_speed: 0 - 108


###Weather report one works perfectly OK
```{r}
weather_report <- function(weather){
  
  #put in checks on data first
  weather %>% 
    verify(temp >= -35 & temp <= 55) %>%            #Set range for temp  
    verify(wind_speed >= 0 & wind_speed <= 108) %>% #Set range for w_s
    verify(humidity >= 0 & humidity <= 100)         #Set range for hum 
  
  # Our average temperature, humidity and wind speed 
  weather_means <- 
  weather %>% 
  summarise(
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_windspeed = mean(wind_speed)
  )

# Show hot and humid days
hot_and_humid <- 
weather %>% 
  filter(temp > 20) %>% 
  filter(humidity > 55)

return(list(weather_means, hot_and_humid))
}

weather_report(weather_1)
```

###Weather report 2 fails as assertr stops execution
code stops after first failure
```{r}
weather_report <- function(weather){
  
  #put in checks on data first
  weather %>% 
    verify(temp >= -35 & temp <= 55) %>%            #Set range for temp  
    verify(wind_speed >= 0 & wind_speed <= 108) %>% #Set range for w_s
    verify(humidity >= 0 & humidity <= 100)         #Set range for hum 
  
  # Our average temperature, humidity and wind speed 
  weather_means <- 
  weather %>% 
  summarise(
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_windspeed = mean(wind_speed)
  )

# Show hot and humid days
hot_and_humid <- 
weather %>% 
  filter(temp > 20) %>% 
  filter(humidity > 55)

return(list(weather_means, hot_and_humid))
}

weather_report(weather_2)
```


### stopifnot Base R function

```{r}
weather_report <- function(weather){
  
  stopifnot( #If these following conditions are NOT TRUE, return error
    nrow(weather) > 15,
    ncol(weather) == 4,
    #check for missing values
    drop_na(weather) == weather
  )
  
  
  
  #put in checks on data first
  weather %>% 
    verify(temp >= -35 & temp <= 55) %>%            #Set range for temp  
    verify(wind_speed >= 0 & wind_speed <= 108) %>% #Set range for w_s
    verify(humidity >= 0 & humidity <= 100)         #Set range for hum 
  
  # Our average temperature, humidity and wind speed 
  weather_means <- 
  weather %>% 
  summarise(
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_windspeed = mean(wind_speed)
  )

# Show hot and humid days
hot_and_humid <- 
weather %>% 
  filter(temp > 20) %>% 
  filter(humidity > 55)

return(list(weather_means, hot_and_humid))
}

weather_report(weather_2)
```




