---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
#Leap-year function ----
is_leap_year <- function (year) {
  if (year %% 100 == 0) {
    if (year %% 400 == 0) {
      return(TRUE)
    } else {
      return(FALSE)
    }
  }
  
  if (year %% 4 == 0) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

is_leap_year("year")
```
###Leap year function explanation

If `if (year %% 4 == 0)` was first in the function, then anything divisible by 4 would be automatically returned TRUE and the function would stop.
Therefore the special centenary years that aren't leap years would be returned TRUE.

Example: year 300 is NOT a leap year and is divisible by 4.

### testthat Package

```{r}
library(testthat)
```

### expect_equal():
checking that the objects give the answer we expect
expect_equal(data, expected return)

### expect_error():
checking thst code throws and error

```{r}
#expect_equal function ----
a <- 10

expect_equal(a, 10)
```

```{r}
expect_equal(a, 11)
```

### test_that(description, {code})

```{r}
test_that("Real leap year returns TRUE", {
  expect_equal(is_leap_year(2020), TRUE)
  expect_equal(is_leap_year(2000), TRUE)
})
```

```{r}
#Task ---
# second test : one expectation
test_that("Character input returns an error", {
  expect_error(is_leap_year("year"))
})
## Test passed 🎉
```

Test name -        "Character input returns an error"
Expectation -       if the string "year" is input to the function, an                        error would be returned
My expectatation -  The test would pass because "year" returns an error                      in the is_leap_year function 


###Test with LOGICAL input
test below fails, as input of TRUE is processed as a '1' and is put through the function
```{r}
test_that("Non-numeeric input returns an error", {
  expect_error(is_leap_year("year"))
  expect_error(is_leap_year(TRUE)) 
})
```

```{r}
#Leap-year function updated ----
is_leap_year <- function (year) {
  
  #add in condition for non-numeric data
  if (!is.numeric(year)) { #This filters inputs and stops function if                               the data is non-numeric
    stop ("Non-numeric input :(")
  }
  
  if (year %% 100 == 0) {
    if (year %% 400 == 0) {
      return(TRUE)
    } else {
      return(FALSE)
    }
  }
  
  if (year %% 4 == 0) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

is_leap_year("year")
```


The test below now passes because TRUE is now non-numeric and throws an error
```{r}
test_that("Non-numeeric input returns an error", {
  expect_error(is_leap_year("year"))
  expect_error(is_leap_year(TRUE)) 
})
```


```{r}
#Task ----
test_that("Test that 2017 and 1900 are non-leap years",{
  expect_equal(is_leap_year(2017), FALSE)
  expect_equal(is_leap_year(1900), FALSE)
})
```

