---
title: "R Notebook"
output: html_notebook
---

What does our distribution look like???

How far apart are the values typically???

Is it a very narrow & tall distribution, or is it more wide & short???

Range

= max value - min value

```{r}
library(tidyverse)
library(janitor)
jobs <- read_csv("data/TyrellCorpJobs.csv") %>% 
  select(-1) %>% 
  clean_names()
```


```{r}
jobs %>% 
  summarise(range = diff(range(salary)))

jobs %>% 
  summarise(range = range(salary))
```
The range of our data is $61,568

> We MUST visualise our data 

```{r}
jobs %>% 
  ggplot(aes(x = salary)) +
  geom_histogram(col = "white", bins = 25)
```

```{r}
jobs %>% 
  ggplot(aes(x = salary)) +
  geom_histogram(col = "white", bins = 25) +
  facet_wrap(~ position)
```
There is strong evidence of __bimodality__


```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    range = diff(range(salary))
  )
```
There is _more_ salary disparity in management than in accounting

Range is also _heavily affected_ by extreme values
Most accounting salaries are between 30-40k - the range suggests a spreak of 15k ~50% more

Less influenced by outliers is the __INTERQUARTILE RANGE__

There are _3_ quartiles which split the data into _4_

> 0 - 25 | _Q1_ | 25 - 100
> 0 - 50 | _Q2_ | 50 - 100 (median)
> 0 - 75 | _Q3_ | 75 - 100

interquartile range = range of q1 to q3

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    q1 = quantile(salary, 0.25),
    q2 = quantile(salary, 0.5),
    q3 = quantile(salary, 0.75),
    iqr = q3 - q1,
    iqr_from_stats = IQR(salary)
  )
```

The five figure summary
   
Min.   :29459 ---> min value
1st Qu.:33038 ---> q1
Median :35009 ---> q2 / median
Mean   :39383 ---> mean
3rd Qu.:41990 ---> q3
Max.   :91027 ---> max value

```{r}
summary(jobs)
```

```{r}
library(skimr)
```



```{r}
jobs %>% 
  group_by(position) %>% 
  skim()
```

Five figues summary using `skim()`

p0   ---> min value
p25  ---> q1
p50  ---> q2 / median
mean ---> mean
p75  ---> q3
p100 ---> max value


A more graphical view of quartiles: Box Plots

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    q1 = quantile(salary, 0.25),
    q2 = quantile(salary, 0.5),
    q3 = quantile(salary, 0.75),
    iqr = q3 - q1,
    iqr_from_stats = IQR(salary),
    outlier_distance = 1.5 * iqr
  )
```
For Accounting:
- IF a value is 4311.75 below q1, this is an outlier 
- IF a value is 4311.75 above q3, this is an outlier

For Management:
- IF a value is 20385.75 below q1, this is an outlier 
- IF a value is 20385.75 above q3, this is an outlier

```{r}
jobs %>% 
  ggplot(aes(x = salary,
             y = position)) +
  geom_boxplot() +
  stat_boxplot(geom = "errorbar", width = 0.5)
```



Task - 10 mins
Investigate and comment on the centrality and spreads of `distribution_1` and `distribution_2` produced by the following function calls.
Use `geom_boxplot()` and `skim()` first
Next, plot histograms to confirm your descriptions.
[Don’t worry about what the functions generating distribution_1 and distribution_2 are doing]

```{r}
set.seed(42) # means we'll all get the same randomly generated data
distribution_1 <- tibble(
  y = append(
    rnorm(n = 100, mean = 5, sd = 10), 
    rnorm(n = 200, mean = 5, sd = 1))
)

distribution_2 <- tibble(
  y = runif(n = 1000, min = -30, max = 30)
)
```

```{r}
distribution_1 %>% 
  ggplot(aes(x = y)) +
  geom_boxplot() +
  stat_boxplot(geom = "errorbar", width = 0.25)
```


```{r}
distribution_1 %>%
  skim()
```



```{r}
distribution_2 %>% 
  ggplot(aes(x = y)) +
  geom_boxplot()  +
  stat_boxplot(geom = "errorbar", width = 0.25)
```


```{r}
distribution_2 %>%
  skim()
```


```{r}
distribution_1 %>% 
  ggplot(aes(x = y)) +
  geom_histogram(col = "white", bins = 50)
```


```{r}
distribution_2 %>% 
  ggplot(aes(x = y)) +
  geom_histogram(col = "white", bins = 50)
```


```{r}
library(e1071)
```


```{r}
distribution_2 %>% 
  summarise(skew = skewness(y, type = 1))
```
highly symmetric distribution


```{r}
right_skewed <- read_csv("data/heavily_right_skewed.csv")
```


```{r}
right_skewed %>% 
  summarise(skew = skewness(x, type = 1))
```
highly right skewed


```{r}
right_skewed %>% 
  ggplot(aes(x = x)) +
  geom_boxplot()
```
> 1. Long tail of outliers to the right of the box
> 2. Q2 (median) is closer to Q1 than to Q3
> 3. Left whisker is shorter than right whisker 


```{r}
right_skewed %>% 
  ggplot(aes(x = x)) +
  geom_histogram(col = "white")
```


```{r}
right_skewed %>% 
  skim()
```

 
 
 ## Variance
 
 > a single number value for a measure of spread
 > a measure of how far each value in the data set is from the mean
 
 [3, 6, 4, 20, 4]
 
```{r}
num_set <- c(3, 6, 4, 20, 4)

mean(num_set)
```
 > sum of the differences from the mean squared and then divided by the total number of elements - 1
 > mean difference from the mean 
 
 A hypothetical example
 
 - a survey people about their weight
  - marathon runners
  - commuters
  - the mean weights are the same

sampled 50 from each group
 
```{r}
set.seed(42)

londoner_weights <- tibble(
  marathoners = rnorm(50, mean = 68, sd = 0.5),
  commuters = rnorm(50, mean = 68, sd = 5)
) %>% 
  pivot_longer(cols = everything(), names_to = "group", values_to = "weight")

ggplot(londoner_weights) +
  aes(x = weight, y = group) +
  geom_boxplot()
```
 
 
```{r}
londoner_weights %>% 
  group_by(group) %>% 
  summarise(
    mean_weight = mean(weight),
    var_weight = var(weight),
    st_dev_weight = sd(weight),
    st_dev_long = sqrt(var_weight)
  )
```
 weight = kg
 variance = kg^2
 std dev = sqrt(varriance)
 std dev = kg
 
 the average deviance from the mean weight of commuters (68.5kg) was 4.6kg
 
 # Calculating variance longhand
 
 $$
 s^2 = \frac{1}{n-1}\sum{(x - \bar{x})^2}
 $$
To calculate variance:
> get the difference between each value and the mean
> square each value
> add up these squared values
> divide by number of total values - 1

To calculate the standard deviation:
> sqrt the variance

Task:
Calculate the variance and standard deviation for the salaries in accounting & management

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(var_salary = var(salary),
            st_dev_salary = sd(salary))
  
```
- The standard deviation of management is > 4x more than that of accounting
- Values of management salaries are much more spread


















