---
title: "R Notebook"
output: html_notebook
---

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(janitor)
```


```{r}
thrones <- read_csv("data/character_data_S01-S08.csv")
```


```{r}
clean_thrones<- thrones %>% 
# Only keep variables of interest
  select(c(sex, religion, occupation, social_status, allegiance_last, allegiance_switched, dth_flag, featured_episode_count, prominence)) %>% 
# Convert to factor level
    mutate(sex = factor(sex, levels = c(1, 2, 9), labels = c("Male", "Female", "Unknown")),
    religion = factor(religion, levels = c(0, 1, 3, 4, 5, 6, 7, 9), labels = c("Great Stallion", "Lord of Light", "Faith of the Seven", "Old Gods", "Drowned God", "Many Faced God", "Other", "Unknown")),
    occupation = factor(occupation, levels = c(1, 2, 9), labels = c("Silk-collar", "Leather-collar", "Unknown")),
    social_status = factor(social_status, levels = c(1, 2, 9), labels = c("Highborn", "Lowborn", "Unknown")),
    allegiance_last = factor(allegiance_last, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9), labels = c("Stark", "Targaryen", "Night's Watch", "Lannister", "Greyjoy", "Bolton", "Frey", "Other", "Unknown")),
    allegiance_switched = factor(allegiance_switched, levels = c(1, 2, 9), labels = c("No", "Yes", "Unknown")),
    dth_flag = factor(dth_flag, levels = c(0, 1), labels = c("Survived", "Died"))) %>%
# Remove NAs 
 na.omit()
```


```{r}
clean_thrones
```

## train / test split

```{r}
set.seed(19)

n_data <- nrow(clean_thrones)

test_index <- sample(1:n_data, size = n_data * 0.2)

thrones_test <- slice(clean_thrones, test_index)

thrones_train <- slice(clean_thrones, -test_index)
```


```{r}
thrones_test %>% 
  tabyl(dth_flag) 
```


```{r}
thrones_train %>% 
  tabyl(dth_flag) 
```


## Build the model

```{r}
thrones_fit <- rpart(
  formula = dth_flag ~ .,
  data = thrones_train,
  method = "class"
)
```


## Visualise decision tree

```{r}
rpart.plot(thrones_fit,
           yesno = 2,
           fallen.leaves = TRUE,
           faclen = 6,
           digits = 4)
```


```{r}
rpart.plot(thrones_fit,
           yesno = 2,
           fallen.leaves = TRUE,
           digits = 4,
           type = 4,
           extra = 101)
```


```{r}
rpart.rules(thrones_fit,
            cover = TRUE)
```


## Create Predictions
```{r}
library(modelr)
```


```{r}
thrones_test_pred <- thrones_test %>% 
  add_predictions(thrones_fit, type = "class")

thrones_test_pred %>% 
  select(dth_flag, pred)
```


```{r}
thrones_test_pred %>% 
  select(prominence, religion, allegiance_last, featured_episode_count, dth_flag, pred)
```


## Check the performance

```{r}
library(yardstick)
```


```{r}
conf_mat <- thrones_test_pred %>% 
  conf_mat(truth = dth_flag,
           estimate = pred)

conf_mat
```


```{r}
thrones_test_pred %>% 
  accuracy(truth = dth_flag,
           estimate = pred)
```


```{r}
thrones_test_pred %>% 
  sensitivity(truth = dth_flag,
           estimate = pred)
```


```{r}
library(caret)

confusionMatrix(thrones_test_pred$pred,
                thrones_test_pred$dth_flag)
```


## Random Forest

```{r}
library(ranger)
```


```{r}
# how to build a machine learning grid
# crossing(num_trees = 8:10 * 100,
#          mtry = 4:6,
#          min.node.size = 3:5)

rf_classifier <- ranger(dth_flag ~ .,
                        data = thrones_train,
                        importance = "impurity",
                        num.trees = 1000, # how many trees
                        mtry = 2, # how to split the variables
                        min.node.size = 5) # hyperparameters
```


train --> test

train      --> 60%
validation --> 20% for validating hyperparameters
re-train on train + val sets
test       --> 20%

k-fold cross validation

```{r}
sort(importance(rf_classifier), dec = TRUE)
```

## Random Forest Output
```{r}
thrones_test_pred <- thrones_test %>% 
  mutate(pred = predict(rf_classifier, thrones_test)$predictions)

thrones_test_pred %>% 
  select(prominence, religion, allegiance_last, featured_episode_count, dth_flag, pred)
```


```{r}
confusionMatrix(thrones_test_pred$pred,
                thrones_test_pred$dth_flag)
```
THis model is slightly more accurate
The TRUE Deaths has increased from 32 --> 38
The FALSE Deaths has decreased from 10 --> 4


