---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)

books <- read_csv("data/books.csv")

books_tidy <- books %>%
  clean_names() %>%
  filter(!is.na(average_rating)) %>%
  rename(num_pages = number_num_pages) %>%
  glimpse()
```


> Question:
Does the proportion of books in the GoodReads db that _lack text reviews_ differ _significantly_ from 7%?


> Step 1 formulate Hypothesis 

```{r}
books_tidy %>% 
  group_by(text_reviews_count) %>% 
  summarise(prop = n() / nrow(books_tidy)) %>% 
  filter(text_reviews_count == 0)
```


Null Hypothesis

The proportion of books that lack text reviews count is not different to 7%
$$
H_0: π_{no\ reviews} = 7\% 
$$


Alternative Hypothesis

The proportion of books that lack text reviews count is different to 7%
$$
H_1: π_{no\ reviews} \neq 7\%
$$


> Step 2 Calculate statistics

$$
\frac{number\ of\ books\ with\ no\ text\ review}{total\ number\ of\ books}
$$
```{r}
books_tidy_prop <- books_tidy %>% 
  mutate(has_no_reviews = text_reviews_count == 0)

books_tidy_prop
```

```{r}
null_distribution <- books_tidy_prop %>% 
  specify(response = has_no_reviews, success = "TRUE") %>% 
  hypothesise(null = "point", p = 0.07) %>% 
  generate(reps = 5000, type = "draw") %>% 
  calculate(stat = "prop")

null_distribution
```

```{r}
null_distribution %>% 
  visualise(bins = 30)
```

```{r}
observed_stat <- books_tidy_prop %>% 
  specify(response = has_no_reviews, success = "TRUE") %>% 
  calculate(stat = "prop")

observed_stat
```


```{r}
null_distribution %>% 
  visualise(bins = 30) +
  shade_p_value(obs_stat = observed_stat, direction = "both") 
```

```{r}
p_value <- null_distribution %>% 
  get_p_value(obs_stat = observed_stat$stat, direction = "both")

p_value
```

Since our p_value (0.026) < alpha (0.05) we can _reject_ the Null hypothesis.
Therefore, our proprtion of books without a text review is _significantly_ different.



