---
title: "R Notebook"
output: html_notebook
---

# Connecting R to a Database

- 1. Connect R to a Postgre database (omni pool)

### Don't put login credentials on github (or any public place)
### Be careful with how we work with our login details:
    -- host name
    -- username
    -- database
    -- port
    -- password
### 2 main strategies to deal with this:
    -- gitignore
    -- store outside of  git repo

### Load in out secret login credentials to R
```{r}
source("secrets.R")
```

## Connecting to a database
```{r}
library(RPostgres)
```

## Create connection
```{r}
db_connection <- dbConnect(
  drv = Postgres(),
  host = om_host,
  port = om_port,
  dbname = om_atabase,
  user = om_username,
  password = om_password,
  bigint = "numeric"
)

# now that we've created our connection object we don't need our 
# credentials any more - we can remove these from our environment

rm(
  om_username,
  om_password
)
```


## Test the connection
```{r}
dbListTables(conn = db_connection)
```


## Running queries
```{r}
employees_table <- dbGetQuery(conn = db_connection, statement = "SELECT * FROM employees;")
```

```{r}
library(tidyverse)
```


```{r}
#Find all employees in Legal dept.
employees_table %>% 
  filter(department == "Legal")
```

## Using SQL code chunks in R Notebooks
### running a query
```{sql, connection = db_connection}
SELECT *
FROM employees
LIMIT 10;
```
### running a query and capturing the output
### creates/updates an object. In this case, output = "out"
```{sql, connection = db_connection, output.var = "out"}
SELECT *
FROM employees
LIMIT 10;
```

### Can call 'out' in R
```{r}
out
```

```{r}
out %>% 
  select(first_name, last_name, department)
```

## Disconnect from database
```{r}

```

















