---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


```{r}
cc_info <- read_csv("data/code_clan_info.csv")
```


```{r}
#if statements ----

if(this_thing_is_true) {#THEN
  #DO THIS THING HERE
  }
```

```{r}
x <- 10
y <- 15
if (x < y) {#THEN
  print("yes")
}

x <- 100

if (x < y) {#THEN
  print("yes")
}
```

```{r}
#Cannot compare one element to many ----
x <- 10
y <- 5:20
if (x < y) {#THEN
  print("yes")
}

```

```{r}
#Check if x is in y's range ----
x <- 10
y <- 5:20
if (x %in% y) {#THEN
  print("yes")
}
```

Write an if statement that prints ‘Yes that divides by 2’ if a number x is divisible by 2 (hint: use %% 2 == 0 as your expression), try it out for different values of x.

```{r}
#Task ----
x <- 10
y <- 2

if(x %% y == 0){
  print("Yes, that divides by 2")
}
```

```{r}
#Adding an else statement ----
x <- 10
y <- 2

if(x %% y == 0){
  print("Yes, that divides by 2")
} else {
  print("No, that does not divide by 2")
}
```

```{r}
x <- 10
y <- 15:20
if (x %in% y) {#THEN
  print("yes")
} else {
  print("no")
}
```

```{r}
x <- 6
y <- 13

if (x %in% c(3, 4, 5)){
  
                       print("x is in our vector")
  
  if (x == 5){
    
              print("x is exactly 5")
    
  }
  
} else {
  
        print("Start of else statement")
  
  if (y %% x == 0){
    
                   print("x divides y")
  
  } else{
    
         print("x does not divide Y")
  }
  
  print ("End of else statement")
  
}
```

When x is 5 and y is 10
x is in our vector
x is exactly 5

When x is 6 and y is 12
Start of else statement
x divides y
End of else statement

When x is 6 and y is 13
Start of else statement
x does not divide Y
End of else statement

'for'
-Used when we _know_ how many iterations are needed
-Will end after desired number of loops
-Can be _slow_, but likely because of a failure to vectorise

'while'
-Used when we _don't_ know how many iterations are needed
-Will end after condition is met
```{r}
#Loops ----
#Bear in mind that loops can usually be AVOIDED ----
#Try and loop under certain CONDITIONS ----

for(item in v) {
  print(item)
}
```

```{r}
for(item in v) {
  print(item * 2)
}
```

### `for()`
```{r}
#Loop through the days of the week ----
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for(i in weekdays){
  # browser() - you can use this function to cycle through the loops in the console
  phrase <- print(paste("The day is", i)) 
}
```

### `while()`
Sometimes you can get stuck in an infinite loop

```{r}
current_value <- 1 #Initialise your variable first
stop_value <- 10

while(current_value < stop_value){
  current_value <- current_value + 1
  print(paste("The value is", current_value))
}
```

### Alternative ways to index a `for` loop

```{r}
weekdays
len_weekdays <- length(weekdays)

for (i in 1:len_weekdays){
  print(paste("The day is", weekdays[i]))
}
```

### Replacing a vector from a loop ___TERRIBLE PRACTICE___

```{r}
weekdays
len_weekdays <- length(weekdays)

for(i in 1:len_weekdays){
  plural <- paste0(weekdays[i], "s") #creating a new value from the old value, based on the index
  weekdays[i] <- plural #modifying each vec element based on the index
}
weekdays
```

```{r}
m <- matrix(1:12, nrow = 4, ncol = 3)
nrows <- nrow(m)
ncols <- ncol(m)

for (row_i in 1:nrows){
  print(m[row_i, ])
}
```

```{r}
m <- matrix(1:12, nrow = 4, ncol = 3)
nrows <- nrow(m)
ncols <- ncol(m)

for (row_i in 1:nrows) {
  for (col_i in 1:ncols) {
    print(m[row_i, col_i])
  }
}
```

```{r}
m <- matrix(rnorm(25), 5)
num_row <- dim(m)[1]
num_col <- dim(m)[2]

for (row_i in 1:num_row){
  for(col_i in 1:num_col){
    print(m[row_i, col_i])
  }
}
```


```{r}
product_matrix <- matrix(NA, nrow = 5, ncol = 5)
num_row <- dim(product_matrix)[1]
num_col <- dim(product_matrix)[2]

for (row_i in 1:num_row){
  for (col_i in 1:num_col){
    product_matrix[row_i, col_i] <- row_i * col_i #At this position, input the product of num_row * num_col
  }
}

product_matrix
```

```{r}
example_matrix = matrix(1:25, nrow = 5, ncol = 5)
example_matrix 

for (row_index in 1:5){
  for (col_index in 1:5){
    if (row_index == col_index){ #When row and col index are the same, print the number at the position of the matrix
      print(example_matrix[row_index, col_index])
    }
   }
}
```

