---
title: "R Notebook"
output: html_notebook
---
###FUNCTIONS

`nchar()`
`nchar(x, type = "chars", allowNA = FALSE, keepNA = NA)`
```{r}
#Need to make sure to either follow the above order ----
nchar("hello", "chars", TRUE,  NA)

#Or state their position explicitly ----
nchar(type = "chars", allowNA = TRUE, x = "hello", keepNA = NA)
```

`paste()`
`paste (..., sep = " ", collapse = NULL, recycle0 = FALSE)`

```{r}
paste("a", "b", "c", "d", sep = " and then, ")
```

### Writing your own function 

```{r}
function_name <- function() {
  # stuff to do
  # 
  # end with the 'return' value
  return(value)
}
```

```{r}
function_name <- function(par_1, par_2 = NULL, par_3 = "default_value"){
  # stuff to do
  # 
  # end with the 'return' value
  return(value)
}
```

```{r}
multiply_by_5 <- function(x) {
  x * 5
}

multiply_by_5(2)
```

```{r}
multiply_by <- function(x, y){
  x * y
}

multiply_by(6, 9)
```

```{r}
multiply_by <- function(x, y = 3){
  x * y
}

multiply_by(6)
```

```{r}
make_price <- function(x) {
  paste0("£", x)
}

make_price(5)
make_price(123)
make_price(10.25)
```

### Triangular function
$$
T_n = \frac{n(n+1)}{2}
$$

### Create traingular function
```{r}
tri_nos <- function(n) {
  n_plus1 <- n + 1
  
  numerator <- n * n_plus1
  
  denominator <- 2
  
  t_n <- numerator / denominator
  
  t_n
}

tri_nos(4)
```

### Create same function in one line

```{r}
tri_nos_2 <- function(n) {
  n * (n + 1) / 2
}

tri_nos_2(4)
```

### Scope

What can a function see?

```{r}
n <- 8

multiply_by_n <- function(x) {
  x * n
}

multiply_by_n(2)
```


```{r}
multiply_by_n <- function(x, n) {
  answer <- x * n
  if (answer %% 2 == 1) {
    z <- paste("x is odd")
  } else{
      z <- paste("x is even")
  }
  z
}

multiply_by_n(3, 3)
```

### Best practice for writing functions

_good_ --- GIVE NAME WITH VERB
_examples_:
`add_3()`
`format_number()`
`read_data()`

_bad_ --- DESCRIPTIVE NAMES
_examples_
`adder()`
`number_formatter()`
`data_reader()`

__DRRY__ principal
Don't repeat yourself


###Debugging function

```{r}
tri_nos <- function(n) {
  n_plus1 <- n + 1
  
  numerator <- n * n_plus1
  
  denominator <- 2
  
  t_n <- numerator / denominator
  
  t_n
}

debugonce(tri_nos)
tri_nos(4)
```

