---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(tidyr)
```


###Question 1
```{r}
diet <- read_csv("data/dietary-composition-by-country.csv")
country_codes <- read_csv("data/country_codes.csv")
```

```{r}
dim(diet)
dim(country_codes)
```

```{r}
names(diet)
names(country_codes)
```



```{r}
diet <- clean_names(diet)
country_codes <- clean_names(country_codes)
```

```{r}
view(diet)
view(country_codes)
```


###Question 2
```{r}
diet_comp_clean <- diet %>% 
  pivot_longer(c(alcoholic_beverages_fao_2017:wheat_fao_2017),
               names_to = "kcal_source",
               values_to = "avg_daily_kcals")
```


###Question 3
```{r}
food_suffix_remove <- c("_fao_2017")

diet_comp_clean <- diet_comp_clean %>% 
  mutate(kcal_source = str_remove_all(kcal_source, food_suffix_remove),
         kcal_source = str_to_title(kcal_source),
         kcal_source = str_replace_all(kcal_source, "_", " ")) %>% 
  rename("country" = "entity")
```

###Question 4

```{r}
diet_comp_clean %>%
  summarise(across(everything(),~sum(is.na(.x))))
```

###Question 5
```{r}
missing_countries <- diet_comp_clean %>% 
  filter(is.na(code)) %>% 
  distinct(country) %>% 
  pull()

missing_countries
```

###Question 6
```{r}
country_codes %>% 
  filter(country %in% c("United States", "Cape Verde"))
```

###Question 7



```{r}
diet_comp_clean <- diet_comp_clean %>% 
  mutate(code = case_when(
    country == "Cabo Verde" ~replace_na(code, "CPV"),
    country == "United States of America" ~replace_na(code, "USA"),
    TRUE ~code))
```


###Question 7 check
```{r}
diet_comp_clean %>% 
  filter(is.na(code))
```

```{r}
view(diet_comp_clean)
```



### Quesiton 8
```{r}
diet_code_test <- diet_comp_clean %>% 
  inner_join(country_codes, by = c("code" = "alpha_3_code", "country")) %>% 
  mutate(code_check = ifelse(alpha_2_code == str_extract_all(code, "[A-Z][A-Z]"),
         TRUE, FALSE))

diet_code_test
```

```{r}
diet_code_test %>% 
  filter(code_check == FALSE) %>% 
  distinct(country)
```

###Question 9 & 10
```{r}
uk_diet_df <- diet_comp_clean %>% 
  filter(country == "United Kingdom",
         !is.na(avg_daily_kcals)) %>%
  mutate(uk_diet = avg_daily_kcals,
         uk_diet_annual_diff = (avg_daily_kcals * 365) - lag(avg_daily_kcals * 365))

uk_diet_df %>% 
  slice_max(uk_diet_annual_diff, n = 1)

uk_diet_df %>% 
  slice_min(uk_diet_annual_diff, n = 1)
```

###Question 11
```{r}
diet_comp_clean %>% 
  filter(year >= 1990,
         kcal_source == "Alcoholic beverages",
         !is.na(avg_daily_kcals)) %>% 
  group_by(year) %>% 
  slice_max(avg_daily_kcals)
```

###Question 12
This was my original code which worked
This code was limited as it only worked to find one country and one year at a time
```{r}

top_calorie_source <- function(countries, years) {
  diet_comp_clean %>% 
  filter(country == countries,
         year == years) %>% 
    slice_max(avg_daily_kcals, n = 1)
}

top_calorie_source("Canada", 1995)
```
UK top calorie source:
1965 - Wheat, 676
1975 - Wheat, 621
1985 - Wheat, 643
1995 - Wheat, 658

Zimbabwe top calorie source:
1965 - Maize, 900
1975 - Maize, 970
1985 - Maize, 971
1995 - Maize, 915

Limitations:
You can only search for a specific country and year, one at a time

###Question 12 Updated
Updates to the above so that the function takes a vector input for both countries and years:
```{r}
top_kcal_source <- function(data, countries, years) {
  diet_comp_clean %>% 
    filter(country %in% countries,
           year %in% years) %>% 
    group_by(country, year) %>% 
    slice_max(avg_daily_kcals, n = 1)
}

top_kcal_source(diet_comp_clean, countries = c("United Kingdom", "Zimbabwe"), years = c(1965, 1975, 1985, 1995))
```


###Question 13

```{r}
b_countries <- diet_comp_clean %>% 
  filter(str_detect(country, "^B")) %>%   #Filter for countries beginning with B
  distinct(country) %>%                   #Use distinct to find individual countries
  pull()                                  #Use pull to put them into a vector

b_countries
```

```{r}
top_kcal_source(diet_comp_clean, b_countries, 1970)
```




