---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


```{r}
source("prob.R")
```

```{r}
s_two_dice <- rolldie(2)

s_two_dice
```

Create event to find possible outcomes for a combination of 7
Use filter function to check totals
```{r}
s_two_dice %>% 
  filter(X1 + X2 == 7)
```

Add new column to store total of dice numbers
```{r}
s_two_dice <- s_two_dice %>% 
  mutate(total = X1 + X2)

s_two_dice
```

```{r}
s_sum_equals_seven <- s_two_dice %>% 
  filter(total == 7)

s_sum_equals_seven
```

```{r}
s_sum_gt_ten <- s_two_dice %>% 
  filter(total > 10)

s_sum_gt_ten
```


Task - 5 mins
Add your own random variable to s_two_dice, and then use your new random variable to define an event space. Some suggestions:
random variable: first_die_even, event: first_die_even == TRUE
random variable: die2_minus_die1, event: die2_minus_die1 == -2
random variable: differ_by_two, event: differ_by_two == FALSE
```{r}
even <- c(2, 4, 6)
s_two_dice <- s_two_dice %>% 
  mutate(first_die_even = X1 %in% even) %>% 
  filter(first_die_even)

s_two_dice
```


same as above but using %% 
```{r}
s_two_dice <- s_two_dice %>% 
  mutate(first_die_even_mod = X1 %% 2 == 0) %>% 
  filter(first_die_even)

s_two_dice
```

