---
title: "R Notebook"
output: html_notebook
---

This is normal text

```{r}
#Code goes here ----
```

Code chunks:

Ctrl + option + i

```{r}
2 + 3 
1 + 2
```
Cmd + Shift + Enter -- run whole code chunk
Cmd + Enter         -- run line where cursor is

Call in existing DF "mtcars"
```{r}
mtcars
```

Cmd + Shift + P   -- Open Command Pallet 

Data Types

Categories
 Nominal
 Ordinal
Numeric
 continuous
 discrete

Use c function to create a vector (elements of same 'type')
```{r}
dbl1 <- c(1, 4, 6, 3)
dbl2 <- c(1, 4, 6, 3, "x")
int1 <- c(1L, 4L, 6L, 3L)
dbl1
dbl2
```

Use class function to check type of data in vector
```{r}
class(dbl1) #Continuous values
class(int1) #Discrete values
```

Use typeof function to check type of data in vector
```{r}
typeof(dbl1)
typeof(dbl2)
```

Use factor function to see different categories
```{r}
char1 <- c("a", "b", "a", "c")
char2 <- factor(char1)
char1
char2
typeof(char1)
typeof(char2)
```

R deals with different types of data in a way that is appropriate for each type

Logical type 
```{r}
log1 <- c(TRUE, FALSE, TRUE, FALSE)
typeof(log1)
```


